Structural-temporal neighborhoods of posts to characterize users in online forums

### 2016-03-04 doc/paper.tex
parallel neighborhood census
data.table to store neighborhoods faster

\graphicspath{ {./img/} }
\DeclareMathOperator*{\argmax}{arg\,max}
\journalname{Comput Stat}

\begin{document}
\title{Structural posts neighborhoods to characterize users in online forums}

@@

An interesting question in online communities is that concerning roles. In sociology, roles are generally seen as the set of expected behaviors that are attached  to a position in the community. Extrapoliting the notion of role, some researchers have looked for roles in online forums. Some others have tried to detect the roles and the users who hold that roles.

Roles can also be studied from the macro or the micro perspective. If studied from the macro, we can analyze the number of users, the percentage of replied post, its centrality in the network, and  so forth.

In this paper, we focus on the anamyse of roles at a micro level, and more specifically at the discussion level. We would like to answer the following question: are there different types of users in terms of the kind of conversation they participate in?

The remaining of the paper is as follows. First we will introduce our representation of online discussions. Then we will explain our concept of structural neighborhood. Then we will apply this to the clustering of users and we will analyze the clusters. Finally, we will attemp to study whether these roles have predictive power.

@@

\begin{figure}
	\centering
	\includegraphics[width=0.32\textwidth]{tree1}
	\includegraphics[width=0.32\textwidth]{tree2}
	\includegraphics[width=0.32\textwidth]{tree3}
	\caption{Representation of discussion threads with post tree graphs. Vertex represent posts and edges represent replies between posts.}
	\label{fig:trees}
\end{figure}
\section{Posts graph neighborhoods}

Formally, these local graphs are known as neighborhoods. Let us consider first the most basic notion of neighborhood in order to understand its limitations. Let $p_i$ be a post in the tree graph $G$. The \textit{neighborhood of $p_i$ at radius $r$} $N_G(p_i,r)$ is the induced subgraph of $G$ consisting of all posts $j \in G$ at distance $d_{ij}\leq r$. In order to analyze the type of discussion where a user participates in, we might count the different neighborhoods in which her posts are embedded. However, the number of possible neighborhoods even with radius 1 is unbounded since, although a post can only have a parent, it can have an infinite number of children (see Figure~ \ref{fig:large_neighboorhood}).

In practice, if we create a matrix $M$ where rows correspond to users, columns correspond to neighborhoods and $M_{ij}$ indicates how many times user $i$ has a post in the neighborhood $j$,  we obtain a very sparse matrix. Moreover, many of the columns correspond to similar type of neighborhoods. Think, for instance, of a post with 1 reply, a post with 10 replies and another post with 15 replies. While the type of conversation that embeds the first posts is different from the other two, it does not make sense to say that the former two are categorically different.

\begin{figure}
	\centering
	\includegraphics[width=0.4\textwidth]{large_neighborhood}
	\includegraphics[width=0.4\textwidth]{small_neighborhood}
	\caption{The size of the neighborhood at distance $R$ is unbounded. These two graph represent frequent neighborhoods of a post (red) that replied to the root (white). However, the tree in the left corresponds to a very successful root while the three in the right has not brought the attention of too many users.}
	\label{fig:large_neighboorhood}
\end{figure}

In order to reduce the space of possible neighborhoods we introduce the notion of \textit{temporal neighborhood}. Let $t_i$ the time where post $p_i$ was added to the thread $G$. Let $\tau_{ij}=|t_i-t_j|$ be the \textit{temporal distance} between two posts $p_i$ and $p_j$. The \textit{temporal neighborhood} $N_{G}^{(t)}(i, r, n)$of spacial radius $r$ and order $n$ of a post $p_i$ is the set of $n$ posts in its spacial neighborhood $N_{G}(i, r)$ that are temporally closest to $i$ and that can reach $i$ trough posts that are also in the temporal neighborhood.

@@

\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth]{PCA}
	\includegraphics[width=0.5\textwidth]{whiskers}
	\caption{PCA projection of the clusters found and cluster profile in every dimension.}
	\label{fig:PCA}
\end{figure}

>>>>

\graphicspath{ {./img/} }
\DeclareMathOperator*{\argmax}{arg\,max}
\journalname{}

\begin{document}
\title{Structural-temporal neighborhoods of posts to characterize users in online forums}

@@

An interesting question in online communities is that concerning roles. In sociology, roles are generally seen as the set of expected behaviours that are attached  to a position in the community. Extrapoliting the notion of role, some researchers have looked for roles in online forums. Some others have tried to detect the roles and the users who hold that roles.

Roles can also be studied from the macro or the micro perspective. If studied from the macro, we can analyse the number of users, the percentage of replied post, its centrality in the network, and  so forth.

In this paper, we focus on the analyse of roles at a micro level, and more specifically at the discussion level. We would like to answer the following question: are there different types of users in terms of the kind of conversation they participate in?

Formally, these local graphs are known as neighborhoods.

The remaining of the paper is as follows. First we will introduce our representation of online discussions. Then we will explain our concept of structural neighbourhood. Then we will apply this to the clustering of users and we will analyze the clusters. Finally, we will attemp to study whether these roles have predictive power.

@@

\begin{figure}
	\centering
	\includegraphics[width=0.32\textwidth]{tree1}
	\includegraphics[width=0.32\textwidth]{tree2}
	\includegraphics[width=0.32\textwidth]{tree3}
	\caption{Representation of discussion threads with post tree graphs. Vertex represent posts and edges represent replies between posts.}
	\label{fig:trees}
\end{figure}

\section{Neighborhoods in discussion trees}
In this section, we discuss about the convenience of classic neighbourhood definition in discussion trees. Then we introduce a new definition to overcome some of the limitations of the classic definition when dealing with temporal trees.

\begin{definition}
Given a tree graph $G$, the \textit{structural neighborhood} of distance $d$ of post $i$ is the induced graph composed of all the vertices that are at distance equal or less than $d$ from post $i$
\end{definition}

From this definition a structuarl neighborhood at any distance has an unbounded number of posts. Figure \ref{fig:large_neighboorhood}, for instance, shows a typical case where a post has a large neighborhood even at distance two since a post can only one parent but infinite children (replies).

\subsection{Structural neighborhoods in discussion trees}

\begin{figure}
	\centering
	\includegraphics[width=0.4\textwidth]{large_neighborhood}
	\includegraphics[width=0.4\textwidth]{small_neighborhood}
	\caption{The size of the neighborhood at distance $R$ is unbounded. These two graph represent frequent neighborhoods of a post (red) that replied to the root (white). However, the tree in the left corresponds to a very successful root while the three in the right has not brought the attention of too many users.}
	\label{fig:large_neighboorhood}
\end{figure}
\subsection{Structuro-temporal neighborhoods in discussion trees}
In order to reduce the space of possible neighborhoods we introduce the notion of \textit{temporal neighborhood}. Let $t_i$ the time where post $p_i$ was added to the thread $G$. Let $\tau_{ij}=|t_i-t_j|$ be the \textit{temporal distance} between two posts $p_i$ and $p_j$. The \textit{temporal neighborhood} $N_{G}^{(t)}(i, r, n)$of spacial radius $r$ and order $n$ of a post $p_i$ is the set of $n$ posts in its spacial neighborhood $N_{G}(i, r)$ that are temporally closest to $i$ and that can reach $i$ trough posts that are also in the temporal neighborhood.

@@

\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth]{PCA_nocolor}
	\includegraphics[width=0.5\textwidth]{whiskers}
	\caption{PCA projection of the clusters found and cluster profile in every dimension.}
	\label{fig:PCA}
\end{figure}
